### 2014-09-10 paper/paper.Rmd
fixed plot sizes, wrote a little bit on shortages

# 1. Teachers, computers, and libraries at school make a difference in math scores

```{r shortages, fig.width=8, fig.height=12, warning=FALSE, message=FALSE, echo=FALSE}
school2012.sub = school2012[, c("name", "SCHOOLID", "SC14Q02", "SC14Q07", "SC14Q10")]
student2012.sub = student2012[, c("name", "SCHOOLID", "PV1MATH", "PV1READ", "PV1SCIE")]

df = scores_by_school %>% left_join(school2012.sub)
df$SC14Q02 <- factor(df$SC14Q02, labels=c(0,1,2,3))
df$SC14Q07 <- factor(df$SC14Q07, labels=c(0,1,2,3))
df$SC14Q10 <- factor(df$SC14Q10, labels=c(0,1,2,3))
df.mshort <- df[!is.na(df$SC14Q02),] %>% group_by(name, SC14Q02) %>% summarise(mmath2 = mean(mmath, na.rm=T))
#qplot(SC14Q02, mmath2, data = df.mshort, facets = ~name, 
#      xlab="Math teacher shortage (0=None, 1=Very little, 2=To some extent, 3=A lot)", 
#      ylab="Math Score", geom=c("point","line"), group=name)
df.lshort <- df[!is.na(df$SC14Q10),] %>% group_by(name, SC14Q10) %>% summarise(mmath2 = mean(mmath, na.rm=T))
#qplot(SC14Q10, mmath2, data = df.lshort, facets = ~name, 
#      xlab="Library shortages (0=None, 1=Very little, 2=To some extent, 3=A lot)", 
#      ylab="Math Score", geom=c("point","line"), group=name)
df2 <- summarise(group_by(df[!is.na(df$SC14Q02),], name, SC14Q02), teachers = mean(mmath, na.rm=T))
colnames(df2)[2] <- "levels"
df3 <- summarise(group_by(df[!is.na(df$SC14Q07),], name, SC14Q07), computers = mean(mmath, na.rm=T))
colnames(df3)[2] <- "levels"
df4 <- summarise(group_by(df[!is.na(df$SC14Q10),], name, SC14Q10), libmaterials = mean(mmath, na.rm=T))
colnames(df4)[2] <- "levels"
df.all <- merge(df2, df3)
df.all <- merge(df.all, df4)
df.m <- melt(df.all, id=c("name", "levels"))
colnames(df.m)[4] <- "math"
qplot(levels, math, data = df.m, facets = ~name, 
      xlab="Shortages (0=None, 1=Very little, 2=To some extent, 3=A lot)", 
      ylab="Math Score", geom=c("point","line"), group=variable, colour=variable) +
      scale_colour_manual("Type", values=c("teachers"="#1B9E77", "computers"="#D95F02", "libmaterials"="#7570B3")) +
      theme(legend.position="bottom") 
#ggsave("shortages.pdf", width=14, height=14)
```

```{r shortagesbars, fig.width=4, fig.height=4, warning=FALSE, message=FALSE, echo=FALSE}
qplot(SC14Q02, data=df, geom="histogram", xlab="Teacher shortage", fill=I("#1B9E77"))
#ggsave("teacher-shortages.pdf", width=5, height=5)
qplot(SC14Q07, data=df, geom="histogram", xlab="Computer shortage", fill=I("#D95F02"))
#ggsave("computer-shortages.pdf", width=5, height=5)
qplot(SC14Q10, data=df, geom="histogram", xlab="Library materials shortage", fill=I("#7570B3"))
#ggsave("library-shortages.pdf", width=5, height=5)
```

# 2. The gender gap in math is not universal but the reading gap is, in favor of girls

@@

>>>>

# 1. The gender gap in math is not universal but the reading gap is, in favor of girls

@@

# 2. Teachers, computers, and libraries at school make a difference in math scores

The school table has information about shortages on items viewed as being important for good instruction. The plot below examines at three of these, lack of teachers, lack of computers and shortage of library materials, in relation to math, reading and science scores from the student table, across countries. For most countries scores decrease on average the greater the shortage.

```{r shortages, fig.width=4, fig.height=4, warning=FALSE, message=FALSE, echo=FALSE}
school2012.sub = school2012[, c("name", "SCHOOLID", "SC14Q02", "SC14Q07", "SC14Q10")]
student2012.sub = student2012[, c("name", "SCHOOLID", "PV1MATH", "PV1READ", "PV1SCIE")]

df = scores_by_school %>% left_join(school2012.sub)
df$SC14Q02 <- factor(df$SC14Q02, labels=c(0,1,2,3))
df$SC14Q07 <- factor(df$SC14Q07, labels=c(0,1,2,3))
df$SC14Q10 <- factor(df$SC14Q10, labels=c(0,1,2,3))
df.mshort <- df[!is.na(df$SC14Q02),] %>% group_by(name, SC14Q02) %>% summarise(mmath2 = mean(mmath, na.rm=T))
#qplot(SC14Q02, mmath2, data = df.mshort, facets = ~name, 
#      xlab="Math teacher shortage (0=None, 1=Very little, 2=To some extent, 3=A lot)", 
#      ylab="Math Score", geom=c("point","line"), group=name)
df.lshort <- df[!is.na(df$SC14Q10),] %>% group_by(name, SC14Q10) %>% summarise(mmath2 = mean(mmath, na.rm=T))
#qplot(SC14Q10, mmath2, data = df.lshort, facets = ~name, 
#      xlab="Library shortages (0=None, 1=Very little, 2=To some extent, 3=A lot)", 
#      ylab="Math Score", geom=c("point","line"), group=name)
df2 <- summarise(group_by(df[!is.na(df$SC14Q02),], name, SC14Q02), teachers = mean(mmath, na.rm=T))
colnames(df2)[2] <- "levels"
df3 <- summarise(group_by(df[!is.na(df$SC14Q07),], name, SC14Q07), computers = mean(mmath, na.rm=T))
colnames(df3)[2] <- "levels"
df4 <- summarise(group_by(df[!is.na(df$SC14Q10),], name, SC14Q10), libmaterials = mean(mmath, na.rm=T))
colnames(df4)[2] <- "levels"
df.all <- merge(df2, df3)
df.all <- merge(df.all, df4)
df.m <- melt(df.all, id=c("name", "levels"))
colnames(df.m)[4] <- "math"
qplot(SC14Q02, data=df, geom="histogram", xlab="Teacher shortage", fill=I("#1B9E77"))
#ggsave("teacher-shortages.pdf", width=5, height=5)
qplot(SC14Q07, data=df, geom="histogram", xlab="Computer shortage", fill=I("#D95F02"))
#ggsave("computer-shortages.pdf", width=5, height=5)
qplot(SC14Q10, data=df, geom="histogram", xlab="Library materials shortage", fill=I("#7570B3"))
#ggsave("library-shortages.pdf", width=5, height=5)
#ggsave("shortages.pdf", width=14, height=14)
```

```{r shortagesbars, fig.width=10, fig.height=12, warning=FALSE, message=FALSE, echo=FALSE}
qplot(levels, math, data = df.m, facets = ~name, 
      xlab="Shortages (0=None, 1=Very little, 2=To some extent, 3=A lot)", 
      ylab="Math Score", geom=c("point","line"), group=variable, colour=variable) +
      scale_colour_manual("Type", values=c("teachers"="#1B9E77", "computers"="#D95F02", "libmaterials"="#7570B3")) +
      theme(legend.position="bottom") 
```
